---
title: "RAM Data - B/BMSY statistics and model fitting"
author:
  - Gregory Britten
output:
  pdf_document: default
  #html_notebook: default
---

Load tidyverse package for data wrangling 
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
```

Load latest version of RAM dataset as an rdata file. 
```{r}
load('data/DBdata[asmt][v4.495].RData') #loads 's
cdn_stocks  <- stock %>% filter(primary_country=="Canada")
cdn_data    <- timeseries_values_views %>% filter(stockid%in%cdn_stocks$stockid)
```

# BMSY analysis

Find intersection of stocks
```{r}
stocks    <- unique(cdn_data$stockid)

dd        <- read_excel('data/Cdn_fishery_attributes_current.xlsx', sheet="2022 Fishery attributes")
dd_stocks <- unique(dd$ram_stock_id)

istocks <- intersect(dd_stocks,stocks)
```

Approximate where B/BMSY not available
```{r}
df <- data.frame(stock=character(),bbmsy=character(),approx=numeric())
for(i in 1:length(istocks)){
  dat   <- cdn_data[cdn_data$stockid==istocks[i],]
  bbmsy <- dat$BdivBmsypref %>% na.omit()
  if(length(bbmsy)>0){
    bbmsy <- bbmsy[length(bbmsy)]
    df <- rbind(df, data.frame(stock=istocks[i],bbmsy=bbmsy,approx=0))
  }else{
    B <- dat$TBbest
    bbmsy <- ifelse(sum(!is.na(B))>0, tail(B[!is.na(B)]/(0.5*max(B,na.rm=TRUE)),1), NA)
    df <- rbind(df, data.frame(stock=istocks[i],bbmsy=bbmsy,approx=1))
  }
}
df <- df[complete.cases(df$bbmsy),] %>% .[order(.$bbmsy),] #order according to magnitude of bbmsy
```

```{r}
write.csv(df, 
          row.names=FALSE,
          file='bbmsy.csv')

```

Make a plot
```{r}
plot(1:nrow(df),df$bbmsy,col=ifelse(df$approx==0,'black','red'),pch=19,xaxt='n',bty='n',xlab='',ylab='B/Bmsy')
  abline(h=1,lty=2)
  axis(side=1,at=1:nrow(df),labels=df$stock,las=2,cex.axis=0.3)
  legend('topleft',legend=c("Assessment B/Bmsy","Approximation"),pch=19,col=c('black','red'),bty='n')
```
# Fit model to B/BMSY

Get rid of NAs
```{r}
df_all <- df[complete.cases(df),]
```

Process data according to previous aggregation rules
```{r}
dat_ram <- left_join(df_all,dd,by=c("stock"="ram_stock_id")) %>%
  mutate(dfo_region = ifelse(dfo_region=="Arctic"|dfo_region=="NCR", 'Central', dfo_region),
                    gear       = ifelse(gear=="SOK"|gear=="Hand", "Hand gathered", gear),
                    gear       = ifelse(gear=="Midwater trawl"|gear=="Purse seine", "Pelagic nets", gear),
                    gear       = ifelse(gear=="Longline"|gear=="Rod-and-reel"|gear=="Troll", "Pelagic lines", gear),
                    gear       = ifelse(gear=="Bottom trawl"|gear=="Dredge", "Demersal nets", gear),
                    gear       = ifelse(gear=="Weir"|gear=="Trap","Trap",gear),
                    taxa       = ifelse(taxa=="Anoplopomatidae"|taxa=="Lophiidae"|taxa=="Myxinidae"|taxa=="Osmeridae"|taxa=="Xiphiidae"|taxa=="Scombridae", "Other", taxa)) %>%
  mutate(fishing_vulnerability = scale(fishing_vulnerability),
                    trophic_level         = scale(trophic_level),
                    log10value            = scale(log10(`value_$` + 1)),
                    ref_points            = scale(ref_points)) %>%
  #filter(status!='Uncertain') %>%
  mutate(dfo_region     = factor(dfo_region),
                  gear           = factor(.$gear),
                  msc_status     = factor(.$msc_status),
                  sea_monitoring = factor(.$sea_monitoring), 
                  taxa           = factor(.$taxa),
                  approx         = factor(.$approx)) %>% 
  mutate(sea_monitoring = relevel(sea_monitoring, ref="Partial"),
                  dfo_region = relevel(dfo_region, ref="Quebec"),
                  msc_status = relevel(msc_status, ref="None"),
                  gear = relevel(gear, ref="Pelagic nets"),
                  taxa = relevel(taxa, ref="Pleuronectidae"))

dat_ram$status <- factor(dat_ram$status, levels=c("Critical", "Cautious", "Healthy"), ordered=TRUE)
```

#Fit full model
```{r}
mod_full <- lm(bbmsy ~
                   fishing_vulnerability + 
                   trophic_level + 
                   log10value + 
                   ref_points +   
                   dfo_region + 
                   gear + 
                   taxa + 
                   msc_status + 
                   sea_monitoring+ 
                   approx, data=dat_ram)
```

#Prune model
```{r}
mod_r <- step(mod_full,trace=0,direction='both') #applies backward stepwise selection algorithm
```

#List continuous variable names to separate for plotting purposes
```{r}
cont_var <- c("log10value","ref_points")  
```

Combine parameter estimates into a big table
```{r}
coefs      <- summary(mod_full)$coefficients[,1]
coefs_r    <- summary(mod_r)$coefficients[,1]
ses_r      <- summary(mod_r)$coefficients[,2]

df <- data.frame(var=names(coefs), coefs=coefs) %>%
  left_join(.,data.frame(var=names(coefs_r), coefs_r=coefs_r),by='var') %>%
  left_join(.,data.frame(var=names(coefs_r), ses_r=ses_r),by='var')
```


Add in the reference category for the categorical variable estimates, then re-group together
```{r,tidy=TRUE,tidy.opts=list(width.cutoff=60)}
int_se    <- summary(mod_r)$coefficients[names(coefs_r)=="(Intercept)",2]
dfcat <- df[!(df$var %in% cont_var| df$var %in% c("(Intercept)")),] %>%
  rbind(.,data.frame(var="dfo_regionQuebec",  coefs=0,coefs_r=0,ses_r=int_se)) %>%
  rbind(.,data.frame(var="taxaPleuronectidae",coefs=0,coefs_r=0,ses_r=int_se)) %>%
  rbind(.,data.frame(var="approx0",coefs=0,coefs_r=0,ses_r=int_se)) %>%
  .[order(.$var),] 
```  


Subtract the mean effect from each variable. This will let visualize the relative effect of each category within each variable. I think this the idea we came up with last time... 
```{r}
nms <- c('dfo_region','taxa','approx')
for(i in 1:length(nms)){
  dfcat[grep(nms[i],dfcat$var),c(2,3)] <- 
    apply(dfcat[grep(nms[i],dfcat$var),c(2,3)], 2, function(x){
      return(x - mean(x))
    })
}  
```  

Order the relative effects across categories within each variable
```{r}
ddf <- dfcat[-c(1:nrow(dfcat)),]
for(i in 1:length(nms)){
  dddf <- dfcat[grep(nms[i],dfcat$var),] %>%
    .[order(.$coefs_r),]
  ddf <- rbind(ddf,dddf)
}
ddf <- ddf[complete.cases(ddf$coefs_r),]
```

Clean up category names for plotting
```{r}
ddf$cat <- ddf$var
for(i in 1:length(nms)){
  ddf$cat[grep(nms[i],ddf$cat)] <- gsub(nms[i], '', ddf$cat[grep(nms[i],ddf$cat)])
} 
ddf$cat[ddf$cat==1] <- 'Yes'
ddf$cat[ddf$cat==0] <- 'No'
```

```{r}
write.csv(ddf[,c(1,3,4,5)],file='coefficients_bbmsy.csv')
```


Make a nice plot
```{r,tidy.opts=list(width.cutoff=60),fig.height = 5, fig.width = 7}
nms <- c('dfo_region','taxa','approx')
cols <- c('black','red','dark green')
par(mar=c(6,3,2,2),oma=c(2,2,2,2),cex.axis=0.8)
plot(-999, xlim=c(1,nrow(ddf)), 
     ylim=c(min(ddf$coefs_r-2*ddf$ses_r,na.rm=TRUE),
            max(ddf$coefs_r+2*ddf$ses_r,na.rm=TRUE)),xaxt='n',bty='n',xlab='',ylab='')
for(i in 1:length(nms)){
    is <- grep(nms[i],ddf$var)
    segments(x0=is,x1=is,
             y0=ddf$coefs_r[is]-1.96*ddf$ses_r[is],
             y1=ddf$coefs_r[is]+1.96*ddf$ses_r[is],lty=2)
    points(is,ddf$coefs_r[is],col=cols[i],pch=19)
    axis(side=1,at=is,labels=ddf$cat[is],las=2,cex.axis=0.7,col=cols[i])
}
abline(h=0,lty=2)
legend('topleft',bty='n',
       legend=c('DFO Region','taxa','B/Bmsy Approximation'),
       pch=19,col=c('black','red','dark green'))
```

Subset continuous variables
```{r}
dfcon <- df[df$var %in% cont_var,] %>% na.omit()
```

```{r}
write.csv(dfcon,row.names=FALSE,file='slopes_bbmsy.csv')
```

Plot slopes of pruned model
```{r,tidy=TRUE, tidy.opts=list(width.cutoff=60)}
par(mar=c(6,3,2,2),oma=c(2,2,2,2),cex.axis=0.7)
  plot(x=1:nrow(dfcon), y=dfcon$coefs_r, xaxt='n', bty='n',ylim=c(-0.5,0.5),xlab='',ylab='',pch=19,xlim=c(0.5,2.5))
  segments(x0=1:nrow(dfcon), x1=1:nrow(dfcon), y0=dfcon$coefs_r-2*dfcon$ses_r, y1=dfcon$coefs_r+2*dfcon$ses_r)
  abline(h=0,lty=2)
  mtext(side=2,line=2.5,'Standardized Coefficient')
  axis(side=1,at=1:nrow(dfcon),labels=c("Log Value","Reference Points"),las=2,cex.axis=0.7)
```
Compute deviations
```{r}
dev <- as.numeric(dat_ram$bbmsy) - as.numeric(predict(mod_r))

devdf <- data.frame(stock=dat_ram$stock_id, dev=dev) %>% 
  .[order(.$dev),]
```

Make a plot. Note we do not need bootstrapping here to enhance resolution of differences between stocks. We could apply bootstrapping for confifdence intervals, however.
```{r}
par(mar=c(6,4,6,2))
plot(1:nrow(devdf), devdf$dev, bty='n', 
     xaxt='n', xlab='', ylab='Deviation', ylim=c(-2,2),
     pch=19,cex=0.5)
  #segments(x0=1:nrow(d), x1=1:nrow(d), y0=pred_tab$pred_boot-pred_tab$pred_sd, y1=pred_tab$pred_boot+pred_tab$pred_sd, lty=3)
  axis(1, at=1:nrow(devdf), labels=devdf$stock, las=2, cex.axis=0.3)
  abline(h=0,lty=2)
```
```{r}
write.csv(devdf, 
          row.names=FALSE,
          file='deviance_bbmsy.csv')

```


Boxplots of B/BMSY vs. Oceana status
```{r}
pdf('BBMSYvsOceana.pdf',height=4.5,width=6.5)
plot(dat_ram$bbmsy ~ dat_ram$status, xlab='', ylab='')
mtext(side=2,expression('B/B'['MSY']),line=2.5)
dev.off()

```




